---
title: "Analysis of GPT-4 Annotation Success"
author: "Sean Trott"
date: "2023-06-05"
output:
  # pdf_document: 
  #     fig_caption: yes
  #     keep_md: yes
  #     keep_tex: yes
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    # code_folding: hide
  # pdf_document: 
  word_document:
    toc: yes
---

# Load libraries and setup

```{r include=FALSE}
library(tidyverse)
library(lme4)
library(ggridges)
library(broom.mixed)
library(lmerTest)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 300, fig.format = "pdf")
```


# Introduction

This file contains analyses that identify the overall correlation between GPT-4's ratings and human ratings for different psycholinguistic norms.

# Dataset 1: Iconicity

```{r}
df_gpt = read_csv("../../data/processed/iconicity/iconicity_gpt-4.csv")
nrow(df_gpt)

df_human = read_csv("../../data/raw/iconicity/iconicity.csv")
nrow(df_human)

df_merged = df_human %>%
  inner_join(df_gpt)
nrow(df_merged)

### How correlated?
cor.test(df_merged$rating, df_merged$gpt4_rating)
cor.test(df_merged$rating, df_merged$gpt4_rating, method = "spearman")

```


```{r iconicity, fig.width=4, fig.height=4, dpi=300}
HUMAN_AGREEMENT_ICONICITY = 0.35

df_corr = df_merged %>%
  summarise(r = cor(rating, gpt4_rating, method = "spearman")) %>%
  mutate(dimension = "Iconicity")

df_corr %>%
  ggplot(aes(x = dimension, y = r)) +
  geom_bar(stat = "identity", alpha = .7) +
  scale_y_continuous(limits = c(0, 1)) +
  geom_hline(yintercept = HUMAN_AGREEMENT_ICONICITY, linetype = "dashed",
             color = "steelblue3", size = 1) +
  labs(x = "", y = "Correlation") +
  theme_minimal()

df_merged %>%
  ggplot(aes(x = gpt4_rating)) +
  geom_histogram(alpha = .5, bins = 7)

df_merged %>%
  ggplot(aes(x = rating)) +
  geom_histogram(alpha = .5, bins = 7)

df_merged %>%
  ggplot(aes(x = gpt4_rating, y = rating)) +
  geom_point(alpha = .6) +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "GPT-4 Iconicity Judgment",
       y = "Human Iconicity Judgment")
```

## Error Analysis

```{r}
df_merged = df_merged %>%
  mutate(diff = gpt4_rating - rating,
         z = (gpt4_rating - rating) / rating_sd,
         abs_diff = abs(diff),
         abs_z = abs(z))

df_merged %>%
  ggplot(aes(x = abs_z)) +
  geom_histogram(alpha = .5, bins = 7)



```

### Qualitative

```{r}
dftop20 = df_merged %>%
  arrange(desc(abs_z)) %>%
  head(20)

summary(dftop20$rating)
summary(dftop20$gpt4_rating)

dftop20
```


# Dataset 2: SimLex

```{r}
df_gpt = read_csv("../../data/processed/simlex/simlex_gpt-4.csv")
nrow(df_gpt)

df_human = read_csv("../../data/raw/simlex/simlex.csv")
nrow(df_human)

df_merged = df_human %>%
  inner_join(df_gpt)
nrow(df_merged)

### How correlated?
cor.test(df_merged$SimLex999, df_merged$gpt4_rating)
cor.test(df_merged$SimLex999, df_merged$gpt4_rating, method = "spearman")
```


```{r simlex, fig.width=4, fig.height=4, dpi=300}

HUMAN_AGREEMENT_SIMLEX = 0.78

df_corr = df_merged %>%
  summarise(r = cor(SimLex999, gpt4_rating, method = "spearman")) %>%
  mutate(dimension = "Similarity (SimLex999)")

df_corr %>%
  ggplot(aes(x = dimension, y = r)) +
  geom_bar(stat = "identity", alpha = .7) +
  scale_y_continuous(limits = c(0, 1)) +
  geom_hline(yintercept = HUMAN_AGREEMENT_SIMLEX, , linetype = "dashed",
             color = "steelblue3", size = 1) +
  labs(x = "", y = "Correlation") +
  theme_minimal()

df_merged %>%
  ggplot(aes(x = gpt4_rating, y = SimLex999)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "GPT-4 Similarity Judgment",
       y = "Human Similarity Judgment")

df_merged %>%
  ggplot(aes(x = gpt4_rating, y = SimLex999)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "GPT-4 Similarity Judgment",
       y = "Human Similarity Judgment") +
  facet_wrap(~SimAssoc333)


```


## Error Analysis

```{r simlex_errors, fig.width=6, fig.height=4, dpi=300}
df_merged = df_merged %>%
  mutate(diff = gpt4_rating - SimLex999,
         abs_diff = abs(diff))

df_merged %>%
  ggplot(aes(x = diff)) +
  geom_histogram(alpha = .5, bins = 7)


df_merged %>%
  ggplot(aes(x = SimAssoc333, y = abs_diff)) +
  geom_jitter(alpha = .1, width = .1) +
  stat_summary (fun = function(x){mean(x)},
                fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
                fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
                geom= 'pointrange', 
                position=position_dodge(width=0.95)) +
  theme_minimal() +
  labs(x = "Associated (0 = no, 1 = yes)",
       y = "Absolute error")

df_merged %>%
  ggplot(aes(y = POS, x = abs_diff)) +
  geom_jitter(alpha = .1, width = .1) +
  stat_summary (fun = function(x){mean(x)},
                fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
                fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
                geom= 'pointrange', 
                position=position_dodge(width=0.95)) +
  theme_minimal() +
  labs(y = "Part-of-Speech",
       x = "Absolute Error")

df_merged %>%
  ggplot(aes(y = factor(concQ), x = abs_diff)) +
  geom_jitter(alpha = .1, width = .1) +
  stat_summary (fun = function(x){mean(x)},
                fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
                fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
                geom= 'pointrange', 
                position=position_dodge(width=0.95)) +
  theme_minimal() +
  labs(y = "Concreteness Quartile",
       x = "Absolute Error")

m_full = lm(data = df_merged, abs_diff ~ concQ + POS)
m_conc = lm(data = df_merged, abs_diff ~ concQ)
m_pos = lm(data = df_merged, abs_diff ~ POS)

anova(m_conc, m_full)
anova(m_pos, m_full)
summary(m_full)

```

### Qualitative

```{r}
dftop20 = df_merged %>%
  arrange(desc(abs_diff)) %>%
  head(20)

summary(dftop20$rating)
summary(dftop20$gpt4_rating)

dftop20
```


# Dataset 3: SimVerb

```{r}
df_gpt = read_csv("../../data/processed/simverb/simverb_gpt-4.csv")
nrow(df_gpt)

df_human = read_csv("../../data/raw/simverb/simverb.csv")
nrow(df_human)

df_merged = df_human %>%
  inner_join(df_gpt)
nrow(df_merged)

### How correlated?
cor.test(df_merged$Similarity, df_merged$gpt4_rating)
cor.test(df_merged$Similarity, df_merged$gpt4_rating, method = "spearman")
```


```{r simverb, fig.width=6, fig.height=4, dpi=300}

HUMAN_AGREEMENT_SIMVERB = 0.86

df_corr = df_merged %>%
  summarise(r = cor(Similarity, gpt4_rating, method = "spearman")) %>%
  mutate(Dimension = "Similarity (SimVerb3500)")

df_corr %>%
  ggplot(aes(x = Dimension, y = r)) +
  geom_bar(stat = "identity", alpha = .7) +
  scale_y_continuous(limits = c(0, 1)) +
  geom_hline(yintercept = HUMAN_AGREEMENT_SIMVERB, , linetype = "dashed",
             color = "steelblue3", size = 1) +
  labs(x = "", y = "Correlation") +
  theme_minimal()


df_merged %>%
  ggplot(aes(x = gpt4_rating, y = Similarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "GPT-4 Similarity Judgment",
       y = "Human Similarity Judgment")

df_merged %>%
  ggplot(aes(x = gpt4_rating, y = Similarity)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "GPT-4 Similarity Judgment",
       y = "Human Similarity Judgment") +
  facet_wrap(~Relation)


```


## Error Analysis

```{r simverb_errors, fig.width=6, fig.height=4, dpi=300}
df_merged = df_merged %>%
  mutate(diff = gpt4_rating - Similarity,
         abs_diff = abs(diff)) 

df_merged %>%
  ggplot(aes(x = diff)) +
  geom_histogram(alpha = .5, bins = 7)


df_merged %>%
  ggplot(aes(y = reorder(Relation, abs_diff), x = abs_diff)) +
  geom_jitter(alpha = .1, width = .1) +
  stat_summary (fun = function(x){mean(x)},
                fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
                fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
                geom= 'pointrange', 
                position=position_dodge(width=0.95)) +
  theme_minimal() +
  labs(y = "Relation Type",
       x = "Absolute Error")

df_merged %>%
  group_by(Relation) %>%
  summarise(m_diff = mean(diff),
            sd_diff = sd(diff))

m = lm(data = df_merged, abs_diff ~ Relation)
m_reduced = lm(data = df_merged, abs_diff ~ 1)
anova(m_reduced, m)
summary(m)
```



### Qualitative

```{r}
dftop20 = df_merged %>%
  arrange(desc(abs_diff)) %>%
  head(20)

summary(dftop20$rating)
summary(dftop20$gpt4_rating)

dftop20
```


# Dataset 4: RAW-C

```{r}
df_gpt = read_csv("../../data/processed/raw-c/raw-c_gpt-4.csv")
nrow(df_gpt)

### How correlated?
cor.test(df_gpt$mean_relatedness, df_gpt$gpt4_rating)
cor.test(df_gpt$mean_relatedness, df_gpt$gpt4_rating, method = "spearman")

```


```{r rawc, fig.width=4, fig.height=4, dpi=300}

HUMAN_AGREEMENT_RAWC = 0.79

df_corr = df_gpt %>%
  summarise(r = cor(mean_relatedness, gpt4_rating, method = "spearman")) %>%
  mutate(Dimension = "Relatedness (RAW-C)")

df_corr %>%
  ggplot(aes(x = Dimension, y = r)) +
  geom_bar(stat = "identity", alpha = .7) +
  scale_y_continuous(limits = c(0, 1)) +
  geom_hline(yintercept = HUMAN_AGREEMENT_RAWC, , linetype = "dashed",
             color = "steelblue3", size = 1) +
  labs(x = "", y = "Correlation") +
  theme_minimal()


df_gpt %>%
  ggplot(aes(x = gpt4_rating, y = mean_relatedness)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "GPT-4 Relatedness Judgment",
       y = "Human Relatedness Judgment")


```


## Error Analysis

```{r rawc_errors, fig.width=6, fig.height=4, dpi=300}
df_gpt = df_gpt %>%
  mutate(diff = gpt4_rating - mean_relatedness,
         abs_diff = abs(diff)) 

df_gpt %>%
  ggplot(aes(x = abs_diff)) +
  geom_histogram(alpha = .5, bins = 7)

df_gpt %>%
  ggplot(aes(y = reorder(same, abs_diff), x = abs_diff)) +
  geom_jitter(alpha = .1, width = .1) +
  stat_summary (fun = function(x){mean(x)},
                fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
                fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
                geom= 'pointrange', 
                position=position_dodge(width=0.95)) +
  theme_minimal() +
  labs(y = "Same Sense",
       x = "Absolute Error")


df_gpt %>%
  ggplot(aes(x = abs_diff,
             y = reorder(ambiguity_type, abs_diff),
             fill = same)) +
  geom_density_ridges2(aes(height = ..density..), 
                       color=gray(0.25), 
                       alpha = 0.5, 
                       scale=0.85, 
                       size=.9, 
                       stat="density") +
  labs(x = "Absolute error",
       y = "Ambiguity Type",
       fill = "Same vs. Different Sense") +
  theme(
    legend.position = "bottom"
  ) + 
  theme(axis.title = element_text(size=rel(1.5)),
        axis.text = element_text(size = rel(1.5)),
        legend.text = element_text(size = rel(1.5)),
        legend.title = element_text(size = rel(1.5)),
        strip.text.x = element_text(size = rel(1.5))) +
  theme_minimal()


summary(lm(data = df_gpt, abs_diff ~ same))

### Get residuals
mod = lm(data = df_gpt, mean_relatedness ~ gpt4_rating + same)
summary(mod)


### Does GPT-4 predict human relatedness for same sense judgments?
mod = lm(data = filter(df_gpt, same == TRUE), mean_relatedness ~ gpt4_rating)
summary(mod)

```


### Qualitative

```{r}

dftop20 = df_gpt %>%
  arrange(desc(abs_diff)) %>%
  head(20)

summary(dftop20$rating)
summary(dftop20$gpt4_rating)

dftop20 %>%
  select(word, sentence1, sentence2, mean_relatedness, gpt4_rating)
```



# Dataset 5a: Contextualized Perception Norms

```{r}
df_gpt = read_csv("../../data/processed/cs_norms_perception/cs_norms_perception_gpt-4.csv")
nrow(df_gpt)

df_human = read_csv("../../data/raw/cs_norms_perception/cs_norms_perception.csv")
nrow(df_human)

df_merged = df_human %>%
  inner_join(df_gpt)
nrow(df_merged)

```


```{r cs_perception, fig.width=4, fig.height=4, dpi=300}
### How correlated?

df_summ = df_merged %>%
  summarise(Vision = cor(Vision.M, Vision, method = "spearman"),
            Hearing = cor(Hearing.M, Hearing, method = "spearman"),
            Touch = cor(Touch.M, Touch, method = "spearman"),
            Olfaction = cor(Olfaction.M, Olfaction, method = "spearman"),
            Taste = cor(Taste.M, Taste, method = "spearman"),
            Interoception = cor(Interoception.M, Interoception, method = "spearman"))

df_summ


df_long = df_summ %>%
  pivot_longer(everything(), names_to = "Factor", values_to = "Correlation")

df_long %>%
  ggplot(aes(x = reorder(Factor, Correlation), y = Correlation)) +
  geom_bar(stat = "identity") +
  labs(x = "", y = "Correlation") +
  scale_y_continuous(limits = c(0,1)) +
  theme_minimal()


### calculate overall correlation
### calculate overall correlation
df_human_long = df_human %>%
  select(word, sentence, Vision.M,
                Hearing.M,
                Touch.M,
                Olfaction.M,
                Taste.M,
                Interoception.M) %>%
  pivot_longer(cols = c(Vision.M,
                Hearing.M,
                Touch.M,
                Olfaction.M,
                Taste.M,
                Interoception.M),
               names_to = "Dimension",
               values_to = "Strength_human") %>%
  mutate(Dimension = str_remove(Dimension, "\\.M$"))

df_gpt_long = df_gpt %>%
  select(word, sentence,Vision, Hearing,Touch,Olfaction,Taste,Interoception) %>%
  pivot_longer(cols = c(Vision, Hearing,Touch,Olfaction,Taste,Interoception),
               names_to = "Dimension",
               values_to = "Strength_GPT")

df_merged_long = df_human_long %>%
  inner_join(df_gpt_long, on = c(word, sentence))

cor.test(df_merged_long$Strength_GPT, df_merged_long$Strength_human)
cor.test(df_merged_long$Strength_GPT, df_merged_long$Strength_human, method = "spearman")


df_merged_long %>%
  ggplot(aes(x = Strength_GPT,
             y = Strength_human)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  labs(x = "Strength (GPT rating)",
       y = "Strength (Human rating)") +
  facet_wrap(~Dimension) +
  theme_minimal()


HUMAN_AGREEMENT_CSP = 0.64

df_corr = data.frame(r = cor(df_merged_long$Strength_human, 
                             df_merged_long$Strength_GPT, method = "spearman"),
                     dimension = "Perception Norms")


df_corr %>%
  ggplot(aes(x = dimension, y = r)) +
  geom_bar(stat = "identity", alpha = .7) +
  scale_y_continuous(limits = c(0, 1)) +
  geom_hline(yintercept = HUMAN_AGREEMENT_CSP, , linetype = "dashed",
             color = "steelblue3", size = 1) +
  labs(x = "", y = "Correlation") +
  theme_minimal()


```




# Dataset 5b: Contextualized Action Norms

```{r}
df_gpt = read_csv("../../data/processed/cs_norms_action/cs_norms_action_gpt-4.csv")
nrow(df_gpt)

df_human = read_csv("../../data/raw/cs_norms_action/cs_norms_action.csv")
nrow(df_human)

df_merged = df_human %>%
  inner_join(df_gpt)
nrow(df_merged)
```

```{r cs_action, fig.width=4, fig.height=4, dpi=300}

### How correlated?

df_summ = df_merged %>%
  summarise(Hand_arm = cor(Hand_arm.M, Hand_arm, method = "spearman"),
            Foot_leg = cor(Foot_leg.M, Foot_leg, method = "spearman"),
            Head = cor(Head.M, Head, method = "spearman"),
            Torso = cor(Torso.M, Torso, method = "spearman"),
            Mouth_throat = cor(Mouth_throat.M, Mouth_throat, method = "spearman"))

df_summ

df_long = df_summ %>%
  pivot_longer(everything(), names_to = "Factor", values_to = "Correlation")

df_long %>%
  ggplot(aes(x = reorder(Factor, Correlation), y = Correlation)) +
  geom_bar(stat = "identity") +
  labs(x = "Factor", y = "Correlation") +
  theme_minimal()

### calculate overall correlation
df_human_long = df_human %>%
  select(word, sentence, Mouth_throat.M, 
         Foot_leg.M, Hand_arm.M, Torso.M, Head.M) %>%
  pivot_longer(cols = c(Mouth_throat.M, Foot_leg.M, Hand_arm.M, Torso.M, Head.M),
               names_to = "Dimension",
               values_to = "Strength_human") %>%
  mutate(Dimension = str_remove(Dimension, "\\.M$"))

df_gpt_long = df_gpt %>%
  select(word, sentence, Mouth_throat, Foot_leg, Hand_arm, Torso, Head) %>%
  pivot_longer(cols = c(Mouth_throat, Foot_leg, Hand_arm, Torso, Head),
               names_to = "Dimension",
               values_to = "Strength_GPT")

df_merged_long = df_human_long %>%
  inner_join(df_gpt_long, on = c(word, sentence))

cor.test(df_merged_long$Strength_GPT, df_merged_long$Strength_human)
cor.test(df_merged_long$Strength_GPT, df_merged_long$Strength_human, method = "spearman")


df_merged_long %>%
  ggplot(aes(x = Strength_GPT,
             y = Strength_human)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  labs(x = "Strength (GPT rating)",
       y = "Strength (Human rating)") +
  facet_wrap(~Dimension) +
  theme_minimal()

HUMAN_AGREEMENT_ACTION = 0.5

df_corr = data.frame(r = cor(df_merged_long$Strength_GPT, 
                             df_merged_long$Strength_human, 
                             method = "spearman"),
                     dimension = "Action Norms")


df_corr %>%
  ggplot(aes(x = dimension, y = r)) +
  geom_bar(stat = "identity", alpha = .7) +
  scale_y_continuous(limits = c(0, 1)) +
  geom_hline(yintercept = HUMAN_AGREEMENT_ACTION, , linetype = "dashed",
             color = "steelblue3", size = 1) +
  labs(x = "", y = "Correlation") +
  theme_minimal()
```




# Dataset 6: Glasgow Norms

```{r}
df_gpt = read_csv("../../data/processed/glasgow/glasgow_gpt-4.csv")
nrow(df_gpt)

### Recode AoA
df_gpt = df_gpt %>%
  mutate(AoA2 = case_when(
    AoA <= 2 ~ 1, 
    AoA <= 4 ~ 2,
    AoA <= 6 ~ 3,
    AoA <= 8 ~ 4,
    AoA <= 10 ~ 5,
    AoA <= 12 ~ 6,
    AoA >= 13 ~ 7
  )) %>%
  mutate(AoA = AoA2)

df_human = read_csv("../../data/raw/glasgow/glasgow.csv")
nrow(df_human)

df_merged = df_human %>%
  inner_join(df_gpt) %>%
  drop_na()
nrow(df_merged)
```


```{r glasgow, fig.width=4, fig.height=4, dpi=300}

### How correlated?

df_summ = df_merged %>%
  summarise(Valence = cor(Valence.M, Valence, method = "spearman"),
            Arousal = cor(Arousal.M, Arousal, method = "spearman"),
            Concreteness = cor(Concreteness.M, Concreteness, method = "spearman"),
            Familiarity = cor(Familiarity.M, Familiarity, method = "spearman"),
            Imageability = cor(Imageability.M, Imageability, method = "spearman"),
            Dominance = cor(Dominance.M, Dominance, method = "spearman"),
            AoA = cor(AoA.M, AoA, method = "spearman"),
            Size = cor(Size.M, Size, method = "spearman"),
            Gender = cor(Gender.M, Gender, method = "spearman"))

df_summ

df_long = df_summ %>%
  pivot_longer(everything(), names_to = "Factor", values_to = "Correlation")

df_long %>%
  ggplot(aes(y = reorder(Factor, Correlation), x = Correlation)) +
  geom_bar(stat = "identity") +
  labs(y = "Factor", x = "Correlation") +
  theme_minimal()

### calculate overall correlation
### calculate overall correlation
df_human_long = df_human %>%
  select(word, Valence.M, Arousal.M, Concreteness.M, Familiarity.M, Imageability.M, Dominance.M, AoA.M, Size.M, Gender.M) %>%
  pivot_longer(cols = c(Valence.M, Arousal.M, Concreteness.M, Familiarity.M, Imageability.M, Dominance.M, AoA.M, Size.M, Gender.M),
               names_to = "Dimension",
               values_to = "Rating_human") %>%
  mutate(Dimension = str_remove(Dimension, "\\.M$"))

df_gpt_long = df_gpt %>%
  select(word,Valence, Arousal, Concreteness, Familiarity, Imageability, Dominance, AoA, Size, Gender) %>%
  pivot_longer(cols = c(Valence, Arousal, Concreteness, Familiarity, Imageability, Dominance, AoA, Size, Gender),
               names_to = "Dimension",
               values_to = "Rating_GPT")

df_merged_long = df_human_long %>%
  inner_join(df_gpt_long, on = c(word)) %>%
  drop_na()

cor.test(df_merged_long$Rating_GPT, df_merged_long$Rating_human)
cor.test(df_merged_long$Rating_GPT, df_merged_long$Rating_human, method = "spearman")


df_merged_long %>%
  ggplot(aes(x = Rating_GPT,
             y = Rating_human)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  labs(x = "Rating (GPT)",
       y = "Rating (Human)") +
  facet_wrap(~Dimension) +
  theme_minimal()


df_corr = data.frame(r = cor(df_merged_long$Rating_human,
                             df_merged_long$Rating_GPT, 
                             method = "spearman"),
                     dimension = "Glasgow Norms")


df_corr %>%
  ggplot(aes(x = dimension, y = r)) +
  geom_bar(stat = "identity", alpha = .7) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "", y = "Correlation") +
  theme_minimal()
```


# Supplementary Analyses

## GPT-3.5 Turbo

### RAW-C

```{r}
df_gpt = read_csv("../../data/processed/raw-c/raw-c_gpt-3.5-turbo.csv") %>%
  mutate(gpt_turbo = `gpt3.5-turbo_rating`)
nrow(df_gpt)

df_human = read_csv("../../data/raw/raw-c/raw-c.csv")

df_merged = df_human %>%
  inner_join(df_gpt)
nrow(df_merged)

### How correlated?
cor.test(df_merged$mean_relatedness, df_merged$gpt_turbo)
cor.test(df_merged$mean_relatedness, df_merged$gpt_turbo, method = "spearman")

```


```{r rawc2, fig.width=4, fig.height=4, dpi=300}

HUMAN_AGREEMENT_RAWC = 0.79

df_corr = df_merged %>%
  summarise(r = cor(mean_relatedness, gpt_turbo, method = "spearman")) %>%
  mutate(Dimension = "Relatedness (RAW-C)")

df_corr %>%
  ggplot(aes(x = Dimension, y = r)) +
  geom_bar(stat = "identity", alpha = .7) +
  scale_y_continuous(limits = c(0, 1)) +
  geom_hline(yintercept = HUMAN_AGREEMENT_RAWC, , linetype = "dashed",
             color = "steelblue3", size = 1) +
  labs(x = "", y = "Correlation") +
  theme_minimal()


df_merged %>%
  ggplot(aes(x = gpt_turbo, y = mean_relatedness)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "GPT-3.5 Turbo Relatedness Judgment",
       y = "Human Relatedness Judgment")


```


## Error Analysis

```{r rawc_errors2, fig.width=6, fig.height=4, dpi=300}
df_merged = df_merged %>%
  mutate(diff = gpt_turbo - mean_relatedness,
         abs_diff = abs(diff)) 

df_merged %>%
  ggplot(aes(x = abs_diff)) +
  geom_histogram(alpha = .5, bins = 7)

df_merged %>%
  ggplot(aes(y = reorder(same, abs_diff), x = abs_diff)) +
  geom_jitter(alpha = .1, width = .1) +
  stat_summary (fun = function(x){mean(x)},
                fun.min = function(x){mean(x) - 2*sd(x)/sqrt(length(x))},
                fun.max = function(x){mean(x) + 2*sd(x)/sqrt(length(x))},
                geom= 'pointrange', 
                position=position_dodge(width=0.95)) +
  theme_minimal() +
  labs(y = "Same Sense",
       x = "Absolute Error")


df_merged %>%
  ggplot(aes(x = abs_diff,
             y = reorder(ambiguity_type, abs_diff),
             fill = same)) +
  geom_density_ridges2(aes(height = ..density..), 
                       color=gray(0.25), 
                       alpha = 0.5, 
                       scale=0.85, 
                       size=.9, 
                       stat="density") +
  labs(x = "Absolute error",
       y = "Ambiguity Type",
       fill = "Same vs. Different Sense") +
  theme(
    legend.position = "bottom"
  ) + 
  theme(axis.title = element_text(size=rel(1.5)),
        axis.text = element_text(size = rel(1.5)),
        legend.text = element_text(size = rel(1.5)),
        legend.title = element_text(size = rel(1.5)),
        strip.text.x = element_text(size = rel(1.5))) +
  theme_minimal()


summary(lm(data = df_merged, abs_diff ~ same))

### Get residuals
mod = lm(data = df_merged, mean_relatedness ~ gpt_turbo + same)
summary(mod)


### Does GPT-turbo predict human relatedness for same sense judgments?
mod = lm(data = filter(df_merged, same == TRUE), mean_relatedness ~ gpt_turbo)
summary(mod)

```
